### 2016-08-03 Projects/DelCo_2012/Report/DelCo_2012.tex
Minor changes to reports, updated West event times in propane gas burner report

\vspace*{2\baselineskip}
July 2016}

@@

\large
\flushright{\bf National Institute of Standards and Technology Technical Note XXXX \\
Natl.~Inst.~Stand.~Technol.~Tech.~Note~XX, \pageref{LastPage} pages (July 2016) \\
% http://dx.doi.org/10.6028/NIST.TN.XXXX \\
CODEN: NTNOEF }

>>>>

\vspace*{2\baselineskip}
August 2016}

@@

\large
\flushright{\bf National Institute of Standards and Technology Technical Note XXXX \\
Natl.~Inst.~Stand.~Technol.~Tech.~Note~XX, \pageref{LastPage} pages (August 2016) \\
% http://dx.doi.org/10.6028/NIST.TN.XXXX \\
CODEN: NTNOEF }

### 2016-08-03 Projects/DelCo_2014_2015/Reports/Hose_Stream_Tests/TeX_Files/DelCo_Hose_Stream_Tests.tex
Minor changes to reports, updated West event times in propane gas burner report

Fire suppression tactics using hose streams can affect ventilation in a structure and may impact the movement of smoke and heat through a structure. Seven series of experiments were conducted to study the impact of different hose stream and nozzle movement pattern combinations on air movement within residential scale structures. The experiments studied four different hose streams: a straight stream, narrow fog stream, and a wide fog stream from a combination nozzle and a solid stream from a smooth bore nozzle. The streams were applied from a static, or fixed, position; by moving the hoseline left-to-right across the room in a sweeping motion; and by rotating the hoseline in both the clockwise and counterclockwise directions. Gas velocity was measured at different locations in the structure during the experiments. The wide fog stream caused the most air movement out of any of the tested streams, reaching a maximum velocity of 2.6~m/s (5.8~mph) and maximum air flow rate of approximately 4.2~m$^3$/s (9000~cfm), followed by the narrow fog stream, and then the straight stream and solid stream from the smooth bore nozzle, which both caused approximately the same amount of air movement through the structure. It was discovered that the straight stream only caused air movement through the structure when it was applied in a moving pattern. Furthermore, there were no statistically significant differences in the average measured air velocity between the clockwise and counterclockwise nozzle movement patterns. It was determined that the type of hose stream and manner in which it is applied dictates the extent to which a stream impacts the ventilation of a structure. To better quantify the impact of such air flows on the fire environment, additional experiments need to be conducted with structure fires in a controlled environment.

@@

One of the objectives of the Fire Research Division at the National Institute of Standards and Technology (NIST) is to improve the safety and effectiveness of firefighters through improved knowledge of fire behavior and examination of how different firefighting tactics affect the fire environment. NIST has conducted a significant amount of research examining how ventilation affects the growth and spread of fire within structures and how the air flow to the fire may be controlled to limit or delay the growth of the fire~\cite{madrzykowski2009fire,kerber2009fire}. The studies have resulted in guidance to the fire service regarding ventilation tactics. However, ventilation tactics alone will not extinguish the fire; fire suppression with hose streams is also needed.

@@

There exists a multitude of different tools, methods, and techniques used to apply water for fire suppression using hose streams. A common type of nozzle that is used to apply water during fire suppression in residential structures is an adjustable-pattern spray nozzle, commonly referred to as a combination nozzle~\cite{NFPA_1964}. Most combination nozzles contain an adjustable tip that can be rotated to produce different stream patterns. The types of streams produced by a combination nozzle are often divided into three categories based on the angle of the stream with respect to the nozzle: a straight stream typically produces an angle less than 15\SI{}{\degree}; a narrow fog stream produces an angle between 15\SI{}{\degree} and 45\SI{}{\degree}; and a wide fog stream produces an angle between 45\SI{}{\degree} and 80\SI{}{\degree}~\cite{IFSTA:Essentials_of_FF}. Increasing the angle of the stream allows for a larger area to be covered by the water spray.  However, increasing the angle decreases the distance reached by the stream and its penetration power.

@@

The interior walls of the first floor of each structure were framed with steel studs set to 400~mm (16~in) centers and track and were lined with 13~mm (0.5~in) thick cement board. The walls were composed of 16~mm (0.63~in) Type X gypsum board. Additionally, the ceiling was composed of two layers of 13~mm (0.5~in) thick cement board.
\FloatBarrier

@@

\subsection{Uncertainty}
\label{sec:uncertainty}
There are different components of uncertainty in the length, differential pressure, flow rate, and gas velocity reported here. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means~\cite{Taylor&Kuyatt:1994}. Type B analysis of systematic uncertainties involves estimating the upper (+a) and lower (-a) limits for the quantity in question such that the probability that the value would be in the interval ($\pm$a) is essentially 100\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then, the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95\% confidence interval (2$\sigma$). For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

\subsubsection{Length Measurements}
Each length measurement was taken carefully. Length measurements such as the room dimensions, instrumentation array locations, and fire apparatus (e.g., nozzle, sprinkler, or fan) placement were made with a hand held laser measurement device which has an accuracy of $\pm$6.0~mm (0.25~in) over a range of 0.61~m (2.0~ft) to 15.3~m (50.0~ft)~\cite{StanleyTools}.  However, conditions affecting the measurement, such as levelness of the device, yields an estimated uncertainty of $\pm$0.5\% for measurements in the 2.0~m (6.6~ft) to 10.0~m (32.8~ft) range. Steel measuring tapes with a resolution of $\pm$0.5~mm (0.02~in) were used to locate individual sensors within a measurement array. The steel measuring tapes were manufactured in compliance with NIST Manual 44, which specifies a tolerance of $\pm$1.6~mm (0.06~in) for 9.1~m (30~ft) tapes and $\pm$6.4~mm (0.25~in) for 30.5~m (100~ft) tapes~\cite{Butcher:2012}.

\subsubsection{Bi-Directional Probes}
Bi-directional probes with pressure transducers and single thermocouples were used to measure the gas velocity. Inconel-sheathed, exposed bead, type K thermocouples were co-located with each probe. A gas velocity measurement study examining the doorway flow of pre-flashover compartment fires yielded expanded uncertainty measurements ranging from $\pm0.14$ to $\pm0.22$ for bi-directional probes of similar design~\cite{Bryant:FSJ2009}. The total expanded uncertainty for gas velocity in these experiments is estimated to be $\pm18$\%.

\subsubsection{Water Flow Rate}
Water flow rate was measured with a pressure and flow meter combination shown in Fig.~\ref{fig:flow_meter}. The meter consists of a section of 63.5~mm (2.5~in) cast aluminum pipe with a 0--4.1~MPa (0--600~psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800~lpm (1250~gpm). The pressure transducer and paddlewheel both connect to the battery operated control box where the pressure transducer voltage is converted to a pressure and the paddlewheel pulse count is converted to a volumetric flow rate. The manufacturer reports a $\pm5$\% calibration expanded uncertainty for the flow sensor and $\pm3$\% for the pressure sensor~\cite{Akron:2009}. The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flow rate. The total expanded uncertainty was estimated at $\pm10$\%.

@@

\caption{Average gas velocity and airflow rate [$\pm 18$\%] through interior doorway when flow path was fully established for each monitor experiment}

@@

& 	&	(cfm)	& 	(-1235 $\pm222$)	& & (3082 $\pm555$) &	(3272 $\pm589$) & & (2397 $\pm432$) &	(2413 $\pm434$) 

@@

& 	&	(cfm)	&	(4483 $\pm807$)		& & (3335 $\pm600$) &	(4980 $\pm896$) & & (2726 $\pm491$) &	(2631 $\pm474$)

@@

& 	&	(cfm)	& 	(9812 $\pm1766$)	& & (4385 $\pm789$) &	(4784 $\pm861$) & & (4183 $\pm753$) &	(2738 $\pm493$)

@@

& 	&	(cfm)	&	(6130 $\pm1103$) 	&	\multicolumn{2}{c}{(6884 $\pm1239$)}	& \multicolumn{4}{c}{}

@@

& 	&	(cfm)	&	(6253 $\pm1126$) 	&	\multicolumn{2}{c}{(7466 $\pm1344$)} 	& \multicolumn{4}{c}{}

@@

\begin{table}[!ht]
\caption{Average gas velocity and airflow rate [$\pm 18$\%] through interior doorway when flow path was fully established for Test~5}
\begin{tabular}{lccccc}
\toprule
 &	&	&	\multicolumn{3}{c}{\underline{Room A}}
\\
\textbf{Stream} & \textbf{Value} & \textbf{Units} &
\textbf{Fixed} & \textbf{Clockwise} & \begin{tabular}{@{}c@{}} \textbf{Counter} \\ \textbf{Clockwise} \\ \end{tabular}
\\ \midrule
\multirow{5}{*}{\textit{Straight}} & \multirow{2}{*}{\small Velocity}
& 		 \small{m/s}   	&   0.2 $\pm0.03$	&   0.7 $\pm0.1$	&	 0.8 $\pm0.2$  \\ 
&	&	 \small{(mph)} 	&  (0.4 $\pm0.1$)	&  (1.5 $\pm0.3$)	&	(1.9 $\pm0.3$)
\\~\\
&	\multirow{2}{*}{ \begin{tabular}{@{}c@{}} \small{Airflow} \\ \small{Rate} \end{tabular}} 
&	 	\small{m$^3$/s}	&	 0.3 $\pm0.1$	&  1.2 $\pm0.2$		&	 1.6 $\pm0.3$  	\\
& 	&	\small{(cfm)}	& 	(628 $\pm113$)	& (2584 $\pm465$) 	&	(3291 $\pm592$)
\\~\\
\multirow{5}{*}{\begin{tabular}{@{}c@{}} \textit{Narrow} \\ \textit{Fog} \end{tabular}} & \multirow{2}{*}{\small Velocity}
& 		\small{m/s}   	&	 1.7 $\pm0.3$   	&	 1.8 $\pm0.3$  	&   1.6 $\pm0.3$   \\
&	&	\small{(mph)} 	&	(3.9 $\pm0.7$) 		&	(4.0 $\pm0.7$) 	&  (3.5 $\pm0.6$)
\\~\\
&	\multirow{2}{*}{ \begin{tabular}{@{}c@{}} \small{Airflow} \\ \small{Rate} \end{tabular}} 
&	 	\small{m$^3$/s}	&	 3.2 $\pm0.6$   	&	 3.3 $\pm0.6$		&	 2.9 $\pm0.5$  	\\
& 	&	\small{(cfm)}	& 	(6786 $\pm1222$)	&	(6977 $\pm1256$) 	&	(6228 $\pm1121$)
\\~\\
\multirow{5}{*}{\begin{tabular}{@{}c@{}} \textit{Wide} \\ \textit{Fog} \end{tabular}} & \multirow{2}{*}{\small Velocity} 
&		\small{m/s} 	&	 1.8 $\pm0.3$   	&   2.1 $\pm0.4$  	&   2.2 $\pm0.4$  \\ 
& 	& 	\small{(mph)}	&	(4.1 $\pm0.7$)  	&  (4.8 $\pm0.9$) 	&  (4.9 $\pm0.9$)  
\\ ~\\
&	\multirow{2}{*}{ \begin{tabular}{@{}c@{}} \small{Airflow} \\ \small{Rate} \end{tabular}} 
&	 	\small{m$^3$/s}	&	 3.4 $\pm0.6$   	&  4.0 $\pm0.7$		&	 4.1 $\pm0.7$  	\\
& 	&	\small{(cfm)}	& 	(7283 $\pm1311$)	& (8463 $\pm1523$) 	&	(8669 $\pm1561$) \\
\midrule
&	&	&	\multicolumn{3}{c}{\underline{Room B Ceiling}}
\\
\textbf{Stream} & \textbf{Value} & \textbf{Units} &
\textbf{Fixed} & \textbf{Clockwise} & \begin{tabular}{@{}c@{}} \textbf{Counter} \\ \textbf{Clockwise} \\ \end{tabular}
\\ \midrule
\multirow{5}{*}{\textit{Straight}} & \multirow{2}{*}{\small Velocity}
& 		 \small{m/s}   	&   -0.1 $\pm0.02$    	&   0.6 $\pm0.1$   	&	 0.5 $\pm0.1$  \\ 
&	&	 \small{(mph)} 	&  (-0.2 $\pm0.04$)  	&  (1.4 $\pm0.3$) 	&	(1.2 $\pm0.2$)
\\~\\
&	\multirow{2}{*}{ \begin{tabular}{@{}c@{}} \small{Airflow} \\ \small{Rate} \end{tabular}} 
&	 	\small{m$^3$/s}	&	 -0.2 $\pm0.03$   	&  1.2 $\pm0.2$		&	 1.0 $\pm0.2$  	\\
& 	&	\small{(cfm)}	& 	(-381 $\pm69$)		& (2469 $\pm445$) 	&	(2155 $\pm388$)
\\~\\
\multirow{5}{*}{\begin{tabular}{@{}c@{}} \textit{Narrow} \\ \textit{Fog} \end{tabular}} & \multirow{2}{*}{\small Velocity}
& 		\small{m/s}   	&   0.9 $\pm0.2$   		&   1.5 $\pm0.3$  	&   1.6 $\pm0.3$   \\
&	&	\small{(mph)} 	&  (2.1 $\pm0.4$)  		&  (3.4 $\pm0.6$) 	&  (3.7 $\pm0.7$)
\\~\\
&	\multirow{2}{*}{ \begin{tabular}{@{}c@{}} \small{Airflow} \\ \small{Rate} \end{tabular}} 
&	 	\small{m$^3$/s}	&	 1.8 $\pm0.3$   	&  2.9 $\pm0.5$		&	 3.1 $\pm0.6$  	\\
& 	&	\small{(cfm)}	& 	(3739 $\pm673$)		& (6046 $\pm1088$) 	&	(6479 $\pm1166$)
\\~\\
\multirow{5}{*}{\begin{tabular}{@{}c@{}} \textit{Wide} \\ \textit{Fog} \end{tabular}} & \multirow{2}{*}{\small Velocity} 
&		\small{m/s} 	&   2.1 $\pm0.4$   		&   2.3 $\pm0.4$  	&   2.6 $\pm0.5$  \\ 
& 	& 	\small{(mph)}	&  (4.7 $\pm0.8$)  		&  (5.2 $\pm0.9$) 	&  (5.8 $\pm1.1$)  
\\ ~\\
&	\multirow{2}{*}{ \begin{tabular}{@{}c@{}} \small{Airflow} \\ \small{Rate} \end{tabular}} 
&	 	\small{m$^3$/s}	&	 3.9 $\pm0.7$   	&  4.3 $\pm0.8$		&	 4.8 $\pm0.9$  	\\
& 	&	\small{(cfm)}	& 	(8271 $\pm1489$)	& (9165 $\pm1650$) 	&	(10273 $\pm1849$) \\
\bottomrule
\end{tabular}
\label{table:east_hand_A6_avgs}
\end{table}

@@

\caption{Average gas velocity and airflow rate [$\pm18$\%] through interior doorway when flow path was fully established for Tests~6 and 7}

@@

Overall, the water flow experiments discussed in this report reveal that hose streams can affect air movement within a structure. The type of hose stream used and manner in which it is applied dictates the extent to which a stream impacts the ventilation of a structure. These experiments have shown there is a potential for hose streams to affect air flows within a structure during fire scenarios. To better quantify the impact of air flows from hose streams on the fire environment, additional experiments need to be conducted with structure fires in a controlled environment.

>>>>

Fire suppression tactics using hose streams can affect ventilation in a structure and may impact the movement of smoke and heat through a structure. Seven series of experiments were conducted to study the impact of different hose stream and nozzle movement pattern combinations on air movement within residential scale structures. The experiments studied four different hose streams: a straight stream, narrow fog stream, and a wide fog stream from a combination nozzle and a solid stream from a smooth bore nozzle. The streams were applied from a static, or fixed, position; by moving the hoseline left-to-right across the room in a sweeping motion; and by rotating the hoseline in both the clockwise and counterclockwise directions. Gas velocity was measured at different locations in the structure during the experiments. The wide fog stream caused the most air movement out of any of the tested streams, reaching a maximum velocity of 2.6~m/s (5.8~mph) and maximum air flow rate of approximately 4.2~m$^3$/s (9000~cfm), followed by the narrow fog stream, and then the straight stream and solid stream from the smooth bore nozzle, which both caused approximately the same amount of air movement through the structure. The data were consistent with a straight stream only causing air movement through the structure when it was applied in a moving pattern. Furthermore, there were no statistically significant differences in the average measured air velocity between the clockwise and counterclockwise nozzle movement patterns. It was determined that the type of hose stream and manner in which it is applied dictates the extent to which a stream impacts the ventilation of a structure. To better quantify the impact of such air flows on the fire environment, additional experiments need to be conducted with structure fires in a controlled environment.

@@

One of the objectives of the Fire Research Division at the National Institute of Standards and Technology (NIST) is to improve the safety and effectiveness of firefighters through improved knowledge of fire behavior and examination of how different firefighting tactics affect the fire environment. NIST has conducted a significant amount of research examining how ventilation affects the growth and spread of fire within structures and how the air flow to the fire may be controlled to limit or delay the growth of the fire~\cite{madrzykowski2009fire,kerber2009fire}. The studies have provided insight for the fire service regarding ventilation tactics. However, ventilation tactics alone will not extinguish the fire; fire suppression with hose streams is also needed.

@@

There exists a multitude of different tools, methods, and techniques used to apply water for fire suppression using hose streams. A common type of nozzle that is used to apply water during fire suppression in residential structures is an adjustable-pattern spray nozzle, commonly referred to as a combination nozzle~\cite{NFPA_1964}. Most combination nozzles contain an adjustable tip that can be rotated to produce different stream patterns. The types of streams produced by a combination nozzle are often divided into three categories based on the angle of the stream with respect to the centerline of the nozzle: a straight stream typically produces an angle less than 15\SI{}{\degree}; a narrow fog stream produces an angle between 15\SI{}{\degree} and 45\SI{}{\degree}; and a wide fog stream produces an angle between 45\SI{}{\degree} and 80\SI{}{\degree}~\cite{IFSTA:Essentials_of_FF}. Increasing the angle of the stream allows for a larger area to be covered by the water spray.  However, increasing the angle decreases the distance reached by the stream and its penetration power.

@@

The interior walls of the first floor of each structure were framed with steel studs set to 400~mm (16~in) centers and track and were lined with 13~mm (0.5~in) thick cement board. A layer of 16~mm (0.63~in) thick Type X gypsum board covered the cement board. Additionally, the ceiling was composed of two layers of 13~mm (0.5~in) thick cement board.
\FloatBarrier

@@

\subsection{Uncertainty}
\label{sec:uncertainty}
There are different components of uncertainty in the length, differential pressure, flow rate, and gas velocity reported here. Uncertainties are grouped into two categories according to the method used to estimate them. Type A uncertainties are those which are evaluated by statistical methods, and Type B are those which are evaluated by other means~\cite{Taylor&Kuyatt:1994}. Type B analysis of systematic uncertainties involves estimating the upper (+a) and lower (-a) limits for the quantity in question such that the probability that the value would be in the interval ($\pm$a) is essentially 100~\%. After estimating uncertainties by either Type A or B analysis, the uncertainties are combined in quadrature to yield the combined standard uncertainty. Then, the combined standard uncertainty is multiplied by a coverage factor of two, which results in the expanded uncertainty with a 95~\% confidence interval (2$\sigma$). For some of these components, such as the zero and calibration elements, uncertainties are derived from referenced instrument specifications. For other components, referenced research results and past experience with the instruments provided input in the uncertainty determination.

\subsubsection{Length Measurements}
Each length measurement was taken carefully. Length measurements such as the room dimensions, instrumentation array locations, and fire apparatus (e.g., nozzle, sprinkler, or fan) placement were made with a hand held laser measurement device which has an accuracy of $\pm$6.0~mm (0.25~in) over a range of 0.61~m (2.0~ft) to 15.3~m (50.0~ft)~\cite{StanleyTools}.  However, conditions affecting the measurement, such as levelness of the device, yields an estimated uncertainty of $\pm$0.5~\% for measurements in the 2.0~m (6.6~ft) to 10.0~m (32.8~ft) range. Steel measuring tapes with a resolution of $\pm$0.5~mm (0.02~in) were used to locate individual sensors within a measurement array. The steel measuring tapes were manufactured in compliance with NIST Manual 44, which specifies a tolerance of $\pm$1.6~mm (0.06~in) for 9.1~m (30~ft) tapes and $\pm$6.4~mm (0.25~in) for 30.5~m (100~ft) tapes~\cite{Butcher:2012}.

\subsubsection{Bi-Directional Probes}
Bi-directional probes with pressure transducers and single thermocouples were used to measure the gas velocity. Inconel-sheathed, exposed bead, type K thermocouples were co-located with each probe. A gas velocity measurement study examining the doorway flow of pre-flashover compartment fires yielded expanded uncertainty measurements ranging from $\pm0.14$ to $\pm0.22$ for bi-directional probes of similar design~\cite{Bryant:FSJ2009}. The total expanded uncertainty for gas velocity in these experiments is estimated to be $\pm18$~\%.

\subsubsection{Water Flow Rate}
Water flow rate was measured with a pressure and flow meter combination shown in Fig.~\ref{fig:flow_meter}. The meter consisted of a section of 63.5~mm (2.5~in) cast aluminum pipe with a 0 to 4.1~MPa (600~psi) pressure transducer and a paddlewheel type flow sensor with a range of 0 to 4800~lpm (1250~gpm). The pressure transducer and paddlewheel were both connected to the battery operated control box where the pressure transducer voltage was converted to a pressure and the paddlewheel pulse count was converted to a volumetric flow rate. The manufacturer reports a $\pm5$~\% calibration expanded uncertainty for the flow sensor and $\pm3$~\% for the pressure sensor~\cite{Akron:2009}. The pressure transducer was calibrated with a known analog pressure gauge. The flow meter was calibrated by capturing water over time and measuring that mass of water to determine the flow rate. The total expanded uncertainty was estimated at $\pm10$~\%.

@@

\caption{Average gas velocity and airflow rate [$\pm 18$~\%] through interior doorway when flow path was fully established for each monitor experiment}

@@

& 	&	(cfm)	& 	(-1250 $\pm200$)	& & (3100 $\pm550$) &	(3250 $\pm600$) & & (2400 $\pm450$) &	(2400 $\pm450$) 

@@

& 	&	(cfm)	&	(4500 $\pm800$)		& & (3350 $\pm600$) &	(5000 $\pm900$) & & (2750 $\pm500$) &	(2650 $\pm450$)

@@

& 	&	(cfm)	& 	(9800 $\pm1750$)	& & (4400 $\pm800$) &	(4800 $\pm850$) & & (4200 $\pm750$) &	(2750 $\pm500$)

@@

& 	&	(cfm)	&	(6150 $\pm1100$) 	&	\multicolumn{2}{c}{(6900 $\pm1250$)}	& \multicolumn{4}{c}{}

@@

& 	&	(cfm)	&	(6250 $\pm1150$) 	&	\multicolumn{2}{c}{(7450 $\pm1350$)} 	& \multicolumn{4}{c}{}

@@

\begin{table}[!ht]
\caption{Average gas velocity and airflow rate [$\pm 18$~\%] through interior doorway when flow path was fully established for Test~5}
\begin{tabular}{lccccc}
\toprule
 &	&	&	\multicolumn{3}{c}{\underline{Room A}}
\\
\textbf{Stream} & \textbf{Value} & \textbf{Units} &
\textbf{Fixed} & \textbf{Clockwise} & \begin{tabular}{@{}c@{}} \textbf{Counter} \\ \textbf{Clockwise} \\ \end{tabular}
\\ \midrule
\multirow{5}{*}{\textit{Straight}} & \multirow{2}{*}{\small Velocity}
& 		 \small{m/s}   	&   0.2 $\pm0.03$	&   0.7 $\pm0.1$	&	 0.8 $\pm0.2$  \\ 
&	&	 \small{(mph)} 	&  (0.4 $\pm0.1$)	&  (1.5 $\pm0.3$)	&	(1.9 $\pm0.3$)
\\~\\
&	\multirow{2}{*}{ \begin{tabular}{@{}c@{}} \small{Airflow} \\ \small{Rate} \end{tabular}} 
&	 	\small{m$^3$/s}	&	 0.3 $\pm0.1$	&  1.2 $\pm0.2$		&	 1.6 $\pm0.3$  	\\
& 	&	\small{(cfm)}	& 	(628 $\pm113$)	& (2584 $\pm465$) 	&	(3291 $\pm592$)
\\~\\
\multirow{5}{*}{\begin{tabular}{@{}c@{}} \textit{Narrow} \\ \textit{Fog} \end{tabular}} & \multirow{2}{*}{\small Velocity}
& 		\small{m/s}   	&	 1.7 $\pm0.3$   	&	 1.8 $\pm0.3$  	&   1.6 $\pm0.3$   \\
&	&	\small{(mph)} 	&	(3.9 $\pm0.7$) 		&	(4.0 $\pm0.7$) 	&  (3.5 $\pm0.6$)
\\~\\
&	\multirow{2}{*}{ \begin{tabular}{@{}c@{}} \small{Airflow} \\ \small{Rate} \end{tabular}} 
&	 	\small{m$^3$/s}	&	 3.2 $\pm0.6$   	&	 3.3 $\pm0.6$		&	 2.9 $\pm0.5$  	\\
& 	&	\small{(cfm)}	& 	(6786 $\pm1222$)	&	(6977 $\pm1256$) 	&	(6228 $\pm1121$)
\\~\\
\multirow{5}{*}{\begin{tabular}{@{}c@{}} \textit{Wide} \\ \textit{Fog} \end{tabular}} & \multirow{2}{*}{\small Velocity} 
&		\small{m/s} 	&	 1.8 $\pm0.3$   	&   2.1 $\pm0.4$  	&   2.2 $\pm0.4$  \\ 
& 	& 	\small{(mph)}	&	(4.1 $\pm0.7$)  	&  (4.8 $\pm0.9$) 	&  (4.9 $\pm0.9$)  
\\ ~\\
&	\multirow{2}{*}{ \begin{tabular}{@{}c@{}} \small{Airflow} \\ \small{Rate} \end{tabular}} 
&	 	\small{m$^3$/s}	&	 3.4 $\pm0.6$   	&  4.0 $\pm0.7$		&	 4.1 $\pm0.7$  	\\
& 	&	\small{(cfm)}	& 	(7283 $\pm1311$)	& (8463 $\pm1523$) 	&	(8669 $\pm1561$) \\
\midrule
&	&	&	\multicolumn{3}{c}{\underline{Room B Ceiling}}
\\
\textbf{Stream} & \textbf{Value} & \textbf{Units} &
\textbf{Fixed} & \textbf{Clockwise} & \begin{tabular}{@{}c@{}} \textbf{Counter} \\ \textbf{Clockwise} \\ \end{tabular}
\\ \midrule
\multirow{5}{*}{\textit{Straight}} & \multirow{2}{*}{\small Velocity}
& 		 \small{m/s}   	&   -0.1 $\pm0.02$    	&   0.6 $\pm0.1$   	&	 0.5 $\pm0.1$  \\ 
&	&	 \small{(mph)} 	&  (-0.2 $\pm0.04$)  	&  (1.4 $\pm0.3$) 	&	(1.2 $\pm0.2$)
\\~\\
&	\multirow{2}{*}{ \begin{tabular}{@{}c@{}} \small{Airflow} \\ \small{Rate} \end{tabular}} 
&	 	\small{m$^3$/s}	&	 -0.2 $\pm0.03$   	&  1.2 $\pm0.2$		&	 1.0 $\pm0.2$  	\\
& 	&	\small{(cfm)}	& 	(-381 $\pm69$)		& (2469 $\pm445$) 	&	(2155 $\pm388$)
\\~\\
\multirow{5}{*}{\begin{tabular}{@{}c@{}} \textit{Narrow} \\ \textit{Fog} \end{tabular}} & \multirow{2}{*}{\small Velocity}
& 		\small{m/s}   	&   0.9 $\pm0.2$   		&   1.5 $\pm0.3$  	&   1.6 $\pm0.3$   \\
&	&	\small{(mph)} 	&  (2.1 $\pm0.4$)  		&  (3.4 $\pm0.6$) 	&  (3.7 $\pm0.7$)
\\~\\
&	\multirow{2}{*}{ \begin{tabular}{@{}c@{}} \small{Airflow} \\ \small{Rate} \end{tabular}} 
&	 	\small{m$^3$/s}	&	 1.8 $\pm0.3$   	&  2.9 $\pm0.5$		&	 3.1 $\pm0.6$  	\\
& 	&	\small{(cfm)}	& 	(3739 $\pm673$)		& (6046 $\pm1088$) 	&	(6479 $\pm1166$)
\\~\\
\multirow{5}{*}{\begin{tabular}{@{}c@{}} \textit{Wide} \\ \textit{Fog} \end{tabular}} & \multirow{2}{*}{\small Velocity} 
&		\small{m/s} 	&   2.1 $\pm0.4$   		&   2.3 $\pm0.4$  	&   2.6 $\pm0.5$  \\ 
& 	& 	\small{(mph)}	&  (4.7 $\pm0.8$)  		&  (5.2 $\pm0.9$) 	&  (5.8 $\pm1.1$)  
\\ ~\\
&	\multirow{2}{*}{ \begin{tabular}{@{}c@{}} \small{Airflow} \\ \small{Rate} \end{tabular}} 
&	 	\small{m$^3$/s}	&	 3.9 $\pm0.7$   	&  4.3 $\pm0.8$		&	 4.8 $\pm0.9$  	\\
& 	&	\small{(cfm)}	& 	(8271 $\pm1489$)	& (9165 $\pm1650$) 	&	(10273 $\pm1849$) \\
\bottomrule
\end{tabular}
\label{table:east_hand_A6_avgs}
\end{table}

@@

\caption{Average gas velocity and airflow rate [$\pm18$~\%] through interior doorway when flow path was fully established for Tests~6 and 7}

@@

For this limited series of water flow experiments, the data demonstrate that for these flows and configurations, hose streams can affect air movement within a structure. The type of hose stream used and manner in which it is applied dictates the extent to which a stream impacts the ventilation of a structure. These experiments have shown there is a potential for hose streams to affect air flows within a structure during fire scenarios. To better quantify the impact of air flows from hose streams on the fire environment, additional experiments need to be conducted with structure fires in a controlled environment.
